---
title: "Standardize one file"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardize one file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette shows the typical workflow to standardize a Snapshot Safari file.


```{r setup}
library(standardizeSnapshot)
```

## Setup a logger

This is an optional but recommended step. If you want not only to print messages to the console, but also to save them in a file, you can use a logger.

The function `create_logger` allows to create a file in the specified location and setup the logging:

```{r}
logfile <- file.path(tempdir(), "logger.log") # create the logfile
logfile

logger <- create_logger(my_logfile = logfile, 
                        console = FALSE)
```

## Read the files

First, we need to read the data. Here, we will use the `digikam` dataframe, which represents Digikam-like data:

```{r}
data("digikam")
head(digikam, 3)
```

We cound also use a simple `read.csv`.
```{r, eval=FALSE}
# Not run
digikam <- read.csv("path/to/file")
```

## Standardize the file

Then, we standardize the file. The function `standardize_snapshot_df` allows to standardize a single dataframe.

This function has a number of options, but the only ones that are mandatory are:

+ `df`: the dataframe to standardize
+ `standard_df`: the reference dataframe telling the function how to rename the columns.

```{r}
std_df <- standardize_snapshot_df(df = digikam, 
                                  standard_df = standard,
                                  locationID_digikam = "MOK",
                                  logger = logger)
```

Here, we also use 2 optional arguments:

+ `locationID_digikam`: a location code, only useful if the data was processed with Digikam: indeed, the location (reserve) cannot be determined from the dataframe alone.
+ `logger`: a logger created with `create_logger`


By default, if the argument `verbose` is `TRUE`, the function displays the head of the first 8 rows of the file. 

## Write files

The last step is to write the standardized files to a destination. For this, we use the function `write_standardized_df`.

This function has only 2 mandatory arguments:

+ `df`: the file to write
+ `to`: the folder in which the file should be copied.

```{r}
final_folder <- file.path(tempdir(), "data") # the folder in which to copy the file
final_folder

write_standardized_df(df = std_df,
                      to = final_folder,
                      logger = logger)
```

Here, we also use the `logger` argument, which is a logger created with `create_logger`.


The file is now written in the destination.
```{r}
list.files(final_folder)
```


We can check that the logger was filled:

```{r}
readLines(logfile)
```

