---
title: "Standardize one file"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardize one file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette shows the typical workflow to standardize a Snapshot Safari file.


```{r setup}
library(standardizeSnapshot)
```

```{r, echo = FALSE, message = FALSE, results = FALSE}
unlink(file.path(tempdir(), "data_in"), 
       recursive = TRUE) 
unlink(file.path(tempdir(), "data_out"), 
       recursive = TRUE) 
unlink(file.path(tempdir(), "log"), 
       recursive = TRUE) 
```

```{r write_data, echo = FALSE, message = FALSE}
# Prepare data
data(digikam)
digikam_write <- digikam
digikam_write$HierarchicalSubject <- paste0("\"", digikam_write$HierarchicalSubject, "\"")

# Create folder
in_folder <- file.path(tempdir(), "data_in")
dir.create(in_folder)

# Write data
write.csv(digikam_write, 
          file.path(in_folder, "digikam.csv"),
          row.names = FALSE,
          quote = FALSE)
```

## Setup a logger

This is an optional but recommended step. If you want not only to print messages to the console, but also to save them in a file, you can use a logger.

The function `create_logger` allows to create a file in the specified location and setup the logging:

```{r}
logfile <- file.path(tempdir(), "log", "logger.log")
logfile

logger <- create_logger(my_logfile = logfile, 
                        console = FALSE)
```

## Read the file

First, we need to read the data. Here, we read a file that was previously written in `r in_folder` (not shown). This dataframe represents Digikam-like data:

```{r}
in_folder <- file.path(tempdir(), "data_in")
in_folder

df <- read.csv(file.path(in_folder,  "digikam.csv"))
head(df, 3)
```

*NB: this file is the same as the `digikam` dataset included in the package.*

You can reproduce the following results by using:

```{r, eval = FALSE}
data(digikam)
df <- digikam
```


## Standardize the file

Then, we standardize the file. The function `standardize_snapshot_df` allows to standardize a single dataframe.

This function has a number of options, but the only ones that are mandatory are:

+ `df`: the dataframe to standardize
+ `standard_df`: the reference dataframe telling the function how to rename the columns. Here, we use the built-in dataset `standard`.

```{r}
std_df <- standardize_snapshot_df(df = df, 
                                  standard_df = standard,
                                  locationID_digikam = "MOK",
                                  logger = logger)
```

Here, we also use 2 optional arguments:

+ `locationID_digikam`: a location code, only useful if the data was processed with Digikam: indeed, the location (reserve) cannot be determined from the dataframe alone.
+ `logger`: a logger created with `create_logger`


By default, the function displays the head of the first 8 columns of the file along with numerous messages. 

## Write the file

The last step is to write the standardized file to a destination. For this, we use the function `write_standardized_df`.

This function has only 2 mandatory arguments:

+ `df`: the file to write
+ `to`: the folder in which the file should be copied.

```{r}
out_folder <- file.path(tempdir(), "data_out") # the folder in which to copy the file
out_folder

write_standardized_df(df = std_df,
                      to = out_folder,
                      logger = logger)
```

Here, we also use the `logger` argument, which is a logger created with `create_logger`.


The file is now written in the destination.
```{r}
list.files(out_folder)
```


We can check that the logger was filled:

```{r}
list.files(file.path(tempdir(), "log"))
readLines(logfile)
```