---
title: "Standardize multiple files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardize multiple files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette shows the typical workflow to standardize multiple Snapshot Safari files.


```{r setup}
library(standardizeSnapshot)
```

## Setup a logger

This is an optional but recommended step. If you want not only to print messages to the console, but also to save them in a file, you can use a logger.

The function `create_logger` allows to create a file in the specified location and setup the logging:

```{r}
logfile <- file.path(tempdir(), "logger.log") # create the logfile
logfile

logger <- create_logger(my_logfile = logfile, 
                        console = FALSE)
```

## Read the files

First, we need to read the data. Here, we will use three built-in datasets: `digikam`, `traptagger` and `zooniverse` (which are respectively Digikam-, TrapTagger- and Zooniverse-like data).

```{r load_digikam}
data("digikam")
head(digikam, 3)
```

```{r load_traptagger}
data("traptagger")
head(traptagger, 3)
```

```{r load_zooniverse}
data("zooniverse")
head(zooniverse, 3)
```

## Format data manually

The data need to be formatted as a list (which must be named for Digikam data, if we want the `locationID` to work properly).

```{r}
dat_list <- list("MOK" = digikam,
                 "ATH" = traptagger,
                 "APN" = zooniverse)
```

**Warning: this structure of "lists of dataframes" can take-up a lot of space. Therefore, it is better to copy the standardized results to the same list to overwrite older results and save space.**


## Automated formatting 

There is a built-in function to read multiple files into a list in the package: the `read_snapshot_files` function. 

The 2 mandatory arguments to this function are:

+ `input`: a vector of valid paths, that can be files and/or folders
+ `basepath`: the part of the path that should be ignored when copying final files (which will be used to name the list).

```{r, eval=FALSE}
# Not run
dat_list <- read_snapshot_files(input = c("path/to/data/dat1.csv",
                                          "path/to/data/dat2.csv",
                                          "path/to/data/dat3.csv")
                                basepath = "path/to/data/")
```

Another useful argument is `except`, which is a vector of files that should be ignored. This is useful if some of the folder(s) in `input` contains files that should be ignored. A typical content of `except` could be:

```{r}
except <- c("reports_FBIP_format_all_recs",
            "README_FILES_for_TEAMS.txt",
            "Special Projects /*",
            "DHP/DHP+OVE_same_file",
            "KGA/KGA-KHO_together")
```

A logger can also be passed-on to this function with the argument `logger` (to copy messages to a logfile).


## Standardize the files

Then, we standardize the file. The function `standardize_snapshot_list` allows to standardize a list of dataframes.

This function has a number of options, but the only ones that are mandatory are:

+ `df_list`: the list of dataframes to standardize
+ `standard_df`: the reference dataframe telling the function how to rename the columns.

```{r}
dat_list <- standardize_snapshot_list(df_list = dat_list,
                                      standard_df = standard,
                                      logger = logger)
```

Here, we also use the optional argument `logger`, which is a logger created with `create_logger`.


By default, the function displays the head of the first 8 columns of each file along with numerous messages. 

## Write files

The last step is to write the standardized files to a destination. For this, we use the function `write_standardized_list`.

This function has only 2 mandatory arguments:

+ `df_list`: the list of files to write
+ `to`: the folder in which the file should be copied.

```{r}
final_folder <- file.path(tempdir(), "data") # the folder in which to copy the file
final_folder

write_standardized_list(df_list = dat_list,
                        to = final_folder,
                        logger = logger)
```

Here, we also use the `logger` argument, which is a logger created with `create_logger`.


The files are now written in the destination.
```{r}
list.files(final_folder)
```


We can check that the logger was filled:

```{r}
readLines(logfile)
```